Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-02-05 article/04_Discussion.Rmd
changes and improvements in the discussion section

One major question asked in this paper was, *if openly available, bulk radiocarbon data does allow the reconstruction of main trends within the distribution of burial rites in Bronze Age Europe*. This is the case. The narrative in section \ref{time-series-results} is surprisingly similar to the one presented by @hardingEuropeanSocietiesBronze2000a. The following paragraphs contain a short evaluation of the time series in Figure \ref{development_burial_type} and \ref{development_burial_construction} in the light of the archaeological state of research for each region. Despite the general success of this methodological approach, unfortunately the amount and quality of data is still not sufficient for a comprehensive picture. The reasons for this lack of data quality are discussed in section \ref{data-quality-discussion}. \newline

@@

The second important question posed for this paper was, *if these time series constructed from radiocarbon data* (Figure \ref{development_burial_type} and \ref{development_burial_construction}) *allow further analysis concerning cultural and spatial distance*? While the idea behind the first question is rather technical -- reproducing already available information in a quantitative way -- the second one has a larger scope. A quantitative representation of an archaeological narrative should path the way for a more diverse set of methods to be applied to this data and cultural distance is an integral topic of archaeological research.

The expectation for this question was a clear correlation of the "cultural" SED distances for both variables *burial type* and *burial construction* as well in between these variables as with spatial distance. The simulation, though with this implementation and parametrisation for sure rather qualitatively than quantitatively, visualises how a high degree of interaction should translate to a significantly higher degree of cultural similarity across human populations (Figure \ref{simulation}). And as the European Bronze Age is usually depicted as a period of trade, travel and exchange [e.g. @kristiansenRiseBronzeAge2005; @sherrattWhatWouldBronzeAge1993] there was no reason to expect something else than Tobler's famous First Law of Geography: "everything is related to everything else, but near things are more related than distant things." [@toblerComputerMovieSimulating1970].

>>>>

One major question asked in this paper was, *if the main trends in the distribution of burial rites in Bronze Age Europe can be detected in openly available bulk radiocarbon data?*. This is the case. The narrative in section \ref{time-series-results} is surprisingly similar to the one presented by @hardingEuropeanSocietiesBronze2000a. The following paragraphs contain a short evaluation of the time series in Figure \ref{development_burial_type} and \ref{development_burial_construction} in the light of the archaeological state of research for each region.

@@

The second important question posed for this paper was, *if these time series constructed from radiocarbon data* (Figure \ref{development_burial_type} and \ref{development_burial_construction}) *allow further analysis concerning cultural and spatial distance*? That can also be affirmed. While the idea behind the first question was rather technical -- reproducing already available information in a quantitative way -- the second one had a larger scope. The quantitative representation of an archaeological narrative paths the way for a diverse set of methods to be applied and the analysis undertaken here is merely the beginning.

Concerning the question of cultural distance, the expectation based on Cultural transmission theory was a clear correlation of the "cultural" SED distances for both variables *burial type* and *burial construction* with spatial distance. The simulation, though with this implementation and parametrisation for sure rather qualitatively than quantitatively, visualises how a high degree of interaction should translate to a higher degree of cultural similarity across human populations (Figure \ref{simulation}). And as the European Bronze Age is usually depicted as a period of trade, travel and exchange [e.g. @kristiansenRiseBronzeAge2005; @sherrattWhatWouldBronzeAge1993] there was no reason to expect something else than Tobler's famous First Law of Geography: "everything is related to everything else, but near things are more related than distant things." [@toblerComputerMovieSimulating1970].
